---
title: "Water Quality and Methane Emissions at Acton Lake"
author: "J. Beaulieu and S. Waldo"
date: "January 26, 2017"
output:
  html_document: default
  pdf_document:
    includes:
      in_header: header.tex
---
## 1. Background:

2. A map showing the location of the sampled sites
3. A 3D map of the reservoir showing the measurement results for :
    + CH~4~ emissions
    + Chlorophyll a
4. Figures showing how Acton Lake compared to the other 31 reservoirs in the study in terms of:
    + CH~4~ emissions
    + Total phosphorus 
    + Total nitrogen
    + Chlorophyll a
5. Tables summarizing the other measured water quality values at each site within Acton Lake    




```{r setup, results='hide', message=FALSE, warning=FALSE, echo=FALSE}
#myWd <- "C:/Users/JBEAULIE/GitRepository/mulitResSurvey" # for Jake
myWd <- "C:/R_Projects/actonFluxProject" # for Sarah

globalWd <- "L:/Priv/Cin/NRMRL/ReservoirEbullitionStudy/actonEddyCovariance/survey" #location of GIS files
```


```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
#Get the data and libraries

# knit() creates a new session and therefore can't read objects directly from
# workspace.  Run this file outside of project session and load project
# workspace.

#not sure about this:
#is this where the "eqAreaData" comes from?
#got it: run through the GRTS routine, then exit, selecting the option to "Save Workspace", then you can load this info by loading .RData
load(paste(myWd, "/.RData", sep = ""))

library(ggplot2)
library(ggmap)
library(dplyr)
library(knitr)
library(rgdal)
```

\newpage

## 2. Map of Sampled Sites    
We sampled Acton Lake on July 10^th^ - 11^th^, August 31^st^ - September 1^st^, and October 4^th^ - 5^th^ of 2017. The sampling sites were chosen using a generalized random tessellation stratified design ("GRTS"), an approach which combines elements of systematic and random survey designs which allows for the random allocation of sampling sites with maximum  spatial coverage of the reservoir. We used a GPS and geographic information system (GIS) software to locate each sampling site (+/- 30 meters).  
  
  

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE, fig.cap='*Location of the fifteen sample sites within Acton Lake. Satellite image from Google Maps.*'}


# Read and project spatial points dataframe for plotting
actonSitesPlot07 <- readOGR(dsn = globalWd, 
                          layer = "acton07SitesEqAreaDataEC")  # shapefile with data for July

actonSitesPlot08 <- readOGR(dsn = globalWd,
                            layer = "acton08SitesEqAreaDataEC") # shapefile with data for Aug

actonSitesPlot10 <- readOGR(dsn = globalWd,
                            layer = "acton10SitesEqAreaDataEC") # shapefile with data for Oct

actonSites84 <- spTransform(x = actonSitesPlot07, #reproject
                            CRS("+proj=longlat +datum=WGS84")) # projection for google maps

actonSites84@data <- mutate(actonSites84@data, 
                            long=coordinates(actonSites84)[,1], # add long to @data slot
                            lat=coordinates(actonSites84)[,2]) # add lat to @data slot

# Get ggmap
# bbox <- make_bbox(data=actonSites84@data, #defines map extent based on sample site lat/lon
#                   long, lat, f = 0.4) # f is zoom.  Large #, less zoom. tweak for each lake.  
# actonSat <- get_map(location = bbox,
#                     color = "color",
#                     source = "google",
#                     maptype = "satellite") #this is a 12.5 Mb file

# ggmap(actonSat) +
#   ylab("Latitude") +
#   xlab ("Longitude") +
#   geom_point(data=filter(actonSites84@data, EvalStatus == "sampled"), #only main sites
#              aes(x=long, y=lat),
#              size = 2, color = "#00BFC4") + # specify color to be consistent across maps
#   theme(axis.title.y = element_blank(), # Eliminate x-axis title
#         plot.title = element_text(hjust = 0.5)) +  #plot title justify center
#   geom_text(data=filter(actonSites84@data, EvalStatus == "sampled"), #only main sites
#             aes(label=siteID, x=long, y=lat),
#             hjust=1.2, vjust=0, size=3, color = "#00BFC4") +      #alternate color: "#F8766D"
#   coord_equal() +
#   ggtitle("Acton Lake Sample Sites")
```

 

\newpage


## 3.  Within-lake values of methane & carbon dioxide diffusive and ebullitive emissions as well as chlorophyll a concentrations.  

July 10^th^ ebullitive methane emissions  

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Methane ebullitive emissions from July 10^th^ - 11^th^ measured at the 15 sample sites within Acton Lake. The height of each "lollipop" corresponds to the emission rate shown on the vertical z-axis in units of milligrams CH~4~ per square meter of lake surface per hour. The x- and y-axes are oriented east-west and north-south, respectively.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot07@data <- merge(actonSitesPlot07@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake") %>%
                               select(ch4.erate.mg.h, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot07@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="July Ebullitive Methane Emissions",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CH[4]~ebullitive~rate~(mg~ CH[4]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,25),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ch4.erate.mg.h)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#FF0000",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


July 10^th^ diffusive methane emissions

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Methane diffusive emissions from July 10^th^ - 11^th^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot07@data <- merge(actonSitesPlot07@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake") %>%
                               select(ch4.drate.mg.h.best, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot07@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="July Diffusive Methane Emissions",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis =0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CH[4]~diffusive~rate~(mg~ CH[4]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,25),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ch4.drate.mg.h.best)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#FF0000",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


August 31^st^ ebullitive methane emissions 

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Methane ebullitive emissions from August 31^st^ - Sept 1^st^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot08@data <- merge(actonSitesPlot08@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Aug") %>%
                               select(ch4.erate.mg.h, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot08@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="August 31st Ebullitive Methane Emissions",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CH[4]~ebullitive~rate~(mg~ CH[4]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,25),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ch4.erate.mg.h)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#FF0000",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


August 31^st^ diffusive methane emissions

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Methane diffusive emissions from August 31^st^ - Sept 1^st^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot08@data <- merge(actonSitesPlot08@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Aug") %>%
                               select(ch4.drate.mg.h.best, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot08@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="August 31st Diffusive Methane Emissions",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CH[4]~diffusive~rate~(mg~ CH[4]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,25),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ch4.drate.mg.h.best)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#FF0000",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


October 4^th^ volumetric gas emissions

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Gas ebullitive emissions from October 4^th^ - 5^th^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot10@data <- merge(actonSitesPlot10@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Oct") %>%
                               select(ebMlHrM2, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot10@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="October 4th Ebullitive Gas Emissions",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(gas~ebullitive~rate~(mL~ gas~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,50),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ebMlHrM2)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#99CCCC",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


October 4^th^ diffusive methane emissions

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Methane diffusive emissions from October 4^th^ - 5^th^ measured at the 15 sample sites within Acton Lake. *' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot10@data <- merge(actonSitesPlot10@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Oct") %>%
                               select(ch4.drate.mg.h.best, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot10@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="October 4th Diffusive Methane Emissions",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CH[4]~diffusive~rate~(mg~ CH[4]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,25),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ch4.drate.mg.h.best)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#FF0000",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

\newpage

July 10^th^ ebullitive CO~2~ emissions  

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Carbon dioxide ebullitive emissions from July 10^th^ - 11^th^ measured at the 15 sample sites within Acton Lake. The height of each "lollipop" corresponds to the emission rate shown on the vertical z-axis in units of milligrams CH~4~ per square meter of lake surface per hour. The x- and y-axes are oriented east-west and north-south, respectively.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot07@data <- merge(actonSitesPlot07@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake") %>%
                               select(co2.erate.mg.h, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot07@data, {
  co2_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(July~Ebullitive~CO[2]~Emissions),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CO[2]~ebullitive~rate~(mg~ CO[2]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,1),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  co2_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(co2.erate.mg.h)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#0000FF",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


July 10^th^ diffusive CO~2~ uptake

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Carbon dioxide diffusive uptake from July 10^th^ - 11^th^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot07@data <- merge(actonSitesPlot07@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake") %>%
                               select(co2.drate.mg.h.best, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot07@data, {
  co2_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(July~Diffusive~CO[2]~Uptake),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis =0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CO[2]~diffusive~rate~(mg~ CO[2]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,125),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  co2_plot$points3d(x=xcoord, y=ycoord, z=(as.numeric(as.character(co2.drate.mg.h.best))*-1), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#0000FF",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


August 31^st^ ebullitive CO~2~ emissions 

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Carbon dioxide ebullitive emissions from August 31^st^ - Sept 1^st^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot08@data <- merge(actonSitesPlot08@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Aug") %>%
                               select(co2.erate.mg.h, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot08@data, {
  co2_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(August~31^{st}~Ebullitive~CO[2]~Emissions),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CO[2]~ebullitive~rate~(mg~ CO[2]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,1),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  co2_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(co2.erate.mg.h)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#0000FF",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


August 31^st^ diffusive CO~2~ uptake

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Carbon dioxide diffusive emissions from August 31^st^ - Sept 1^st^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot08@data <- merge(actonSitesPlot08@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Aug") %>%
                               select(co2.drate.mg.h.best, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot08@data, {
  co2_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(August~31^{st}~Diffusive~CO[2]~Uptake),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CO[2]~diffusive~rate~(mg~ CO[2]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,125),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  co2_plot$points3d(x=xcoord, y=ycoord, z=(as.numeric(as.character(co2.drate.mg.h.best))*-1), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#0000FF",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


October 4^th^ volumetric gas emissions

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Gas ebullitive emissions from October 4^th^ - 5^th^ measured at the 15 sample sites within Acton Lake.*' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot10@data <- merge(actonSitesPlot10@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Oct") %>%
                               select(ebMlHrM2, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot10@data, {
  ch4_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(October~4^{th}~Ebullitive~Gas~Emissions),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(Gas~ebullitive~rate~(mL~ gas~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,50),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  ch4_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(ebMlHrM2)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#99CCCC",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```


October 4^th^ diffusive CO~2~ uptake

```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Carbon dioxide diffusive emissions from October 4^th^ - 5^th^ measured at the 15 sample sites within Acton Lake. *' }


actonEqArea <- readOGR(dsn = globalWd, # get polygon
                           layer = "actonEqArea",  # shapefile name
                            verbose = FALSE)  #keep from reporting progress

actonSitesPlot10@data <- merge(actonSitesPlot10@data, # add emission to point shp
                             filter(eqAreaData, Lake_Name == "Acton Lake Oct") %>%
                               select(co2.drate.mg.h.best, siteID))

# load external script (this is the scatterplot3d function, but changed a bit with to add more user control over style)
source(paste(myWd, "/scriptsAndRmd/scatterplot3d_edit.R", sep = ""))

# grab coordinates from polygon (in order to plot as a line on the xy plane of 3D plot)
coords = fortify(actonEqArea)
coords <- subset(coords, !hole)



##########################################################################
### Total CH4 emission rate
with(actonSitesPlot10@data, {
  co2_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   
                                 scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(October~4^{th}~Diffusive~CO[2]~Uptake),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                    cex.axis = 0.9, 
                                 # cex.axis = 1.3,
                                   cex.lab = 1.1,
                                 #cex.lab = 2,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=expression(CO[2]~diffusive~rate~(mg~ CO[2]~ m^{-2}~ hr^{-1})),                      # z-axis label
                                   zlim=range(0,125),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   tick.marks = TRUE,
                                  label.tick.marks = TRUE,   #whether or not to include tick mark labels for all three axes
                                 x.ticklabs=NULL,
                                 y.ticklabs=NULL,
                                 
                                 grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
                                    
                                 )
  

  # add the lollipop points
  co2_plot$points3d(x=xcoord, y=ycoord, z=(as.numeric(as.character(co2.drate.mg.h.best))*-1), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#0000FF",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

\newpage

Similarly, chlorophyll a concentrations in Acton Lake did not show a clear difference between upstream and downstream sites.  

Chla July
```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Chlorophyll a concentrations measured at the 15 sample sites within Acton Lake in July. The height of each "lollipop" corresponds to the concentration shown on the vertical z-axis in units of micrograms per liter of water. The x- and y-axes are oriented east-west and north-south, respectively.*' }
##########################################################################
### chlorophyll plot


with(actonSitesPlot07@data, {
  chl_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="July 10th Chl a Levels",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                   cex.axis = 0.9,
                                   cex.lab = 1.1,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=(expression(chlorophyll~a~paste("(", {mu}, "g ", L^{-1}, ")", sep = ""))),                      # z-axis label
                                   zlim=range(0, 45),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
  )
  

  # add the lollipop points
  chl_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(chla_S)), # x y and z data
  #chl_plot$points3d(x=xcoord, y=ycoord, z=chla_num, # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#339966",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

Chla Aug
```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Chlorophyll a concentrations measured at the 15 sample sites within Acton Lake in August. The height of each "lollipop" corresponds to the concentration shown on the vertical z-axis in units of micrograms per liter of water. The x- and y-axes are oriented east-west and north-south, respectively.*' }
##########################################################################
### chlorophyll plot



with(actonSitesPlot08@data, {
  chl_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="August 31st Chl a Levels",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                   cex.axis = 0.9,
                                   cex.lab = 1.1,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=(expression(chlorophyll~a~paste("(", {mu}, "g ", L^{-1}, ")", sep = ""))),                      # z-axis label
                                   zlim=range(0, 45),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
  )
  

  # add the lollipop points
  chl_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(chla_S)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#339966",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

Chla Oct
```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Chlorophyll a concentrations measured at the 15 sample sites within Acton Lake in October. The height of each "lollipop" corresponds to the concentration shown on the vertical z-axis in units of micrograms per liter of water. The x- and y-axes are oriented east-west and north-south, respectively.*' }
##########################################################################
### chlorophyll plot


with(actonSitesPlot10@data, {
  chl_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main="October 4th Chl a Levels",      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                   cex.axis = 0.9,
                                   cex.lab = 1.1,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=(expression(chlorophyll~a~paste("(", {mu}, "g ", L^{-1}, ")", sep = ""))),                      # z-axis label
                                   zlim=range(0, 45),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
  )
  

  # add the lollipop points
  chl_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(chla_S)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#339966",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

\newpage

Turbidity July
```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Chlorophyll a concentrations measured at the 15 sample sites within Acton Lake in July. The height of each "lollipop" corresponds to the concentration shown on the vertical z-axis in units of micrograms per liter of water. The x- and y-axes are oriented east-west and north-south, respectively.*' }
##########################################################################
### turbidity plot


with(actonSitesPlot07@data, {
  turb_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(July~10^{th}~Turbidity),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                   cex.axis = 0.9,
                                   cex.lab = 1.1,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=(expression(turbidity~paste("(", "NTU" ,")", sep = ""))),                      # z-axis label
                                   zlim=range(0, 45),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
  )
  

  # add the lollipop points
  turb_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(TrNTU_S)), # x y and z data
  #chl_plot$points3d(x=xcoord, y=ycoord, z=chla_num, # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#993300",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

Turbidity Aug
```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Chlorophyll a concentrations measured at the 15 sample sites within Acton Lake in August. The height of each "lollipop" corresponds to the concentration shown on the vertical z-axis in units of micrograms per liter of water. The x- and y-axes are oriented east-west and north-south, respectively.*' }
##########################################################################
### turbidity plot



with(actonSitesPlot08@data, {
  turb_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=expression(August~31^{st}~Turbidity),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                   cex.axis = 0.9,
                                   cex.lab = 1.1,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=(expression(turbidity~paste("(", "NTU" ,")", sep = ""))),                      # z-axis label
                                   zlim=range(0, 45),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
  )
  

  # add the lollipop points
  turb_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(TrNTU_S)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#993300",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

Turbidity Oct
```{r echo=FALSE, warning=F, message=F, error=F, fig.width=7, fig.height=7, fig.cap='*Chlorophyll a concentrations measured at the 15 sample sites within Acton Lake in October. The height of each "lollipop" corresponds to the concentration shown on the vertical z-axis in units of micrograms per liter of water. The x- and y-axes are oriented east-west and north-south, respectively.*' }
##########################################################################
### Turbidity plot


with(actonSitesPlot10@data, {
  turb_plot <- scatterplot3d_edit(x=coords$long, y=coords$lat, z=rep(0,length(coords$long)),   # x y and z data, respectively
                                   color="#282830",                               # color of lake outline
                                   lwd=2.0,                                       # line width of lake outline
                                   type="l",                                      # plot type - "l" = line, "h" = the lollipop style with a line  connecting to the horizontal plane 
                                   angle=40,                                      # angle aspect of plot (-40)
                                   scale.y=0.5,                                   # scale y axis (increase by 175% (1.75))
                                   main=(October~4^{th}~Turbidity),      # main plot title
                                   xlab="",                                       # x-axis label (left blank here to reposition later)
                                   ylab="",                                       # y-axis label
                                   cex.axis = 0.9,
                                   cex.lab = 1.1,
                                   mar = c(5,6,4,2)+0.1,
                                   zlab=(expression(turbidity~paste("(", "NTU" ,")", sep = ""))),                      # z-axis label
                                   zlim=range(0, 45),                            # range of z data shown - I grabbed the max from previously plotting the methane data here. Set so when I add the points, the z-axis fits that data. 
                                   lty.axis=1,                                    # the line type of the axes
                                   axis=TRUE,                                     # wheather to include axes
                                   grid=c('xy','xz','yz'),                        # which grids to show - added to edited version of scatterplot3d fx
                                   lty.grid.xy = 1,                               # grid line type  
                                   lty.grid.xz = 3,
                                   lty.grid.yz = 3,
                                   lwd.grid.xy = 1,                               # line width of grid 
                                   lwd.grid.xz = 1,
                                   lwd.grid.yz = 1,
                                   box=FALSE                                      # get rid of the box around the plot
  )
  

  # add the lollipop points
  turb_plot$points3d(x=xcoord, y=ycoord, z=as.numeric(as.character(TrNTU_S)), # x y and z data
                      col="#282830",                      # color of lines connecting points to lake outline
                      lwd=1.5,                            # line width of lines connecting points to lake outline
                      pch=21,                             # type of point to plot
                      bg="#993300",                       # fill color of points
                      type="h",                           # plot type = lines to the horizontal plane
                      cex=2 #(vRateBest/500) + 0.5          # scaling of point sizes - this will need to be adjusted for each variable
  )
})
```

\newpage

\blandscape

## 5. Tables of measured values at each site  

Table 1: Surface (0.1 m) water quality parameters measured via sonde at each site on July 10^th^

```{r warning=FALSE, echo=FALSE}
######################################################################
#### Item #5: Table of Water Quality Values by Site
## work in progress. Goal: produce a table for each lake with water quality and 
## GHG emissions info, including data that was not in the plots. Need to isolate
## the relevant data for each lake, rename the column headings, and add units

# dplyr package 
#"filter" command filters rows of interest. We're filtering the dataset for the 
# sampled sites at a given lake.
#"select" command selects the columns we want to show. The order they are listed 
#in the () is the order they will appear
#used info from http://j-wang.blogspot.com/2015/03/landscape-and-portrait-in-rmarkdown.html to orient table pages as landscape
actonTableSonde <- filter(eqAreaData, Lake_Name == "Acton Lake", 
                      EvalStatus == "sampled") %>%
                select(siteID, LatSamp, LongSmp,  #site identification  
                chla_S, DO__L_S,  pH_S, SpCn__S, Tmp_C_S,  TrNTU_S) #shallow sonde

kable(actonTableSonde,
      align = "c",
      digits = c(1,5,5,1,1,1,0,1,1),
      col.names = c("Site ID", "Latitude", "Longitude",
                    "Chl a (ug/L)", "DO (mg/L)", "pH",
                    "Sp. Cond", "Water Temp (C)", 
                    "Turbidity"))
```


Table 2: Surface (0.1 m) water quality parameters measured via sonde at each site on August 31^st^

```{r warning=FALSE, echo=FALSE}
######################################################################
#### Item #5: Table of Water Quality Values by Site
## work in progress. Goal: produce a table for each lake with water quality and 
## GHG emissions info, including data that was not in the plots. Need to isolate
## the relevant data for each lake, rename the column headings, and add units

# dplyr package 
#"filter" command filters rows of interest. We're filtering the dataset for the 
# sampled sites at a given lake.
#"select" command selects the columns we want to show. The order they are listed 
#in the () is the order they will appear
#used info from http://j-wang.blogspot.com/2015/03/landscape-and-portrait-in-rmarkdown.html to orient table pages as landscape
actonTableSonde <- filter(eqAreaData, Lake_Name == "Acton Lake Aug", 
                      EvalStatus == "sampled") %>%
                select(siteID, LatSamp, LongSmp,  #site identification  
                chla_S, DO__L_S,  pH_S, SpCn__S, Tmp_C_S,  TrNTU_S) #shallow sonde

kable(actonTableSonde,
      align = "c",
      digits = c(1,5,5,1,1,1,0,1,1),
      col.names = c("Site ID", "Latitude", "Longitude",
                    "Chl a (ug/L)", "DO (mg/L)", "pH",
                    "Sp. Cond", "Water Temp (C)", 
                    "Turbidity"))
```


Table 3: Surface (0.1 m) water quality parameters measured via sonde at each site on October 4^th^

```{r warning=FALSE, echo=FALSE}
######################################################################
#### Item #5: Table of Water Quality Values by Site
## work in progress. Goal: produce a table for each lake with water quality and 
## GHG emissions info, including data that was not in the plots. Need to isolate
## the relevant data for each lake, rename the column headings, and add units

# dplyr package 
#"filter" command filters rows of interest. We're filtering the dataset for the 
# sampled sites at a given lake.
#"select" command selects the columns we want to show. The order they are listed 
#in the () is the order they will appear
#used info from http://j-wang.blogspot.com/2015/03/landscape-and-portrait-in-rmarkdown.html to orient table pages as landscape
actonTableSonde <- filter(eqAreaData, Lake_Name == "Acton Lake Oct", 
                      EvalStatus == "sampled") %>%
                select(siteID, LatSamp, LongSmp,  #site identification  
                chla_S, DO__L_S,  pH_S, SpCn__S, Tmp_C_S,  TrNTU_S) #shallow sonde

kable(actonTableSonde,
      align = "c",
      digits = c(1,5,5,1,1,1,0,1,1),
      col.names = c("Site ID", "Latitude", "Longitude",
                    "Chl a (ug/L)", "DO (mg/L)", "pH",
                    "Sp. Cond", "Water Temp (C)", 
                    "Turbidity"))
```


\elandscape

## 6.  EPA disclaimer

The U.S. Environmental Protection Agency, through its Office of Research and Development, participated in the research described herein. It has been subjected to the Agency's administrative review and has been approved for limited external distribution. Any opinions expressed in this article are those of the authors and do not necessarily reflect the views of the Agency, therefore, no official endorsement should be inferred. Any mention of trade names or commercial products does not constitute endorsement or recommendation for use.
